---
title: "ASS_3_clean"
author: "Annamaya Halskov-Jensen"
date: "18/5/2019"
output: html_document
---

```{r setup, include=FALSE} 
setwd("/Users/annamayahalskov-jensen/Documents/Cognitive Science/4th semester/Computational modeling for cognitive science/Multivariate linear models")
 
library(pacman)
p_load("coda", "mvtnorm", "devtools", "rstan", "rethinking", "dplyr", "dfcrm", "brms", "ggplot2")
# library(devtools)
# devtools::install_github("paul-buerkner/brms")
# pacman::p_load(brms)
#update.packages("brms", dependencies = T)
#data
data <- read.csv("Assignment3Data.csv")
```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

### 1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results.
P.S. Should you scale? 
P.P.S. Don't forget to motivate your priors. 
P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.

```{r filtering data}
# only visit 1 
visit1 <- data %>% filter(Visit == 1)
#visit1 <- na.omit(visit1)
# only visit 1  + only asd
v1_asd <- filter(data, data$Visit == 1 & data$ASD == 1)
```

```{r standardising variables}
v1_asd$ADOS = (v1_asd$ADOS-mean(v1_asd$ADOS, na.rm = T)) / sd(v1_asd$ADOS)
v1_asd$VerbalIQ = (v1_asd$VerbalIQ-mean(v1_asd$VerbalIQ, na.rm = T)) / sd(v1_asd$VerbalIQ)
v1_asd$NonVerbalIQ = (v1_asd$NonVerbalIQ-mean(v1_asd$NonVerbalIQ, na.rm = T)) / sd(v1_asd$NonVerbalIQ)
v1_asd$SocialIQ = (v1_asd$SocialIQ-mean(v1_asd$SocialIQ, na.rm = T)) / sd(v1_asd$SocialIQ)
```


DAG???
```{r}
install.packages("dagitty")
library(dagitty)
install.packages("V8")
library("V8")
g <- dagitty('dag {
    ASD_true [pos="0,2.5"]
    Q1 [pos="1,2"]
    Qn [pos="1,3"]
    VIQ [pos="2,4"]
    NVIQ [pos="2,3"]
    SIQ [pos="2,2"]
    ADOS [pos="2,1"]
    ASD_diagnosis [pos="3,2.5"]
    ASD_true -> Q1 -> VIQ -> ASD_diagnosis
    ASD_true -> Q1 -> NVIQ -> ASD_diagnosis
    ASD_true -> Q1 -> SIQ -> ASD_diagnosis
    ASD_true -> Q1 -> ADOS -> ASD_diagnosis
    ASD_true -> Qn -> VIQ -> ASD_diagnosis
    ASD_true -> Qn -> NVIQ -> ASD_diagnosis
    ASD_true -> Qn -> SIQ -> ASD_diagnosis
    ASD_true -> Qn -> ADOS -> ASD_diagnosis
}')
plot(g)
```


1.1. Verbal IQ and ADOS
```{r vIQ + ados} 
plot(v1_asd$ADOS ~ v1_asd$VerbalIQ)   
 
#  defining model first
m0 <-  bf(ADOS ~ 1 + VerbalIQ)
dens (v1_asd$VerbalIQ)
dens (v1_asd$ADOS)
 
#Getting the priors
get_prior(m0, v1_asd, family = gaussian)
#Defining priors used
prior = c(prior(normal(0,0.5), class = 'Intercept'), prior(normal(0,1), class = 'b'), prior(normal(0,0.8), class = 'sigma'))
#modelling
m0_prior <-  brm(m0, family = gaussian, v1_asd, prior = prior, sample_prior = "only")
#Checking priors
pp_check(m0_prior, nsample = 100)
### MAP MAP MAP
#MAP
flist <-  alist(
    ADOS ~ dnorm(mu , sigma),
    mu <- a + b*VerbalIQ ,
        a ~ dnorm( 0 , 0.5 ) ,
        b ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dunif( 0 , 1 )
  )
m.m0 <- rethinking::map(flist, data = v1_asd)
precis(m.m0) 
precis_plot(precis(m.m0))
```

```{r plots} 
### Plot of raw data 
ggplot(v1_asd, aes(VerbalIQ, ADOS)) +
  geom_point()
### Another one
# extract 20 samples from the posterior
post <- extract.samples( m.m0 , n=100 )
# display raw data and sample size
plot( ADOS ~ VerbalIQ , data=v1_asd )
abline( a=coef(m.m0)["a"] , b=coef(m.m0)["b"], col=col.alpha("red", 1))
for (i in 1:150)
abline( a=post$a[i], b=post$b[i], col=col.alpha("blue",0.1))
### New plot
#Look at the distribution of mu
verb.seq <- seq(from= -2, to= 2.5, by=0.1)
#Link-function to compute mu for each sample from posterior and for each value in the sequence
mu <- link( m.m0 , data=data.frame(VerbalIQ=verb.seq) )
str(mu)
# use type="n" to hide raw data
plot( ADOS~ VerbalIQ , v1_asd , type="n" )
# loop over samples and plot each mu value
for ( i in 1:nrow(mu) )
points( verb.seq , mu[i,] , pch=16 , col=col.alpha(rangi2,0.1) )
### GOODPPLOT NICEPLOT
# summarize the distribution of mu
mu.mean <- apply(mu , 2 , mean ) 
mu.HPDI <- apply(mu , 2 , HPDI , prob=0.89 )
# plot raw data
# fading out points to make line and interval more visible
plot( ADOS ~ VerbalIQ , data=v1_asd , col=col.alpha(rangi2,1) )
# plot the MAP line, aka the mean mu for each weight
lines( verb.seq , mu.mean )
# plot a shaded region for 89% HPDI
shade( mu.HPDI , verb.seq )
```


1.2. Non Verbal IQ and ADOS
```{r nIQ + ados}  
plot(v1_asd$ADOS ~ v1_asd$NonVerbalIQ) 
 
#  defining model first
m2 <-  bf(ADOS ~ 1 + NonVerbalIQ)
dens (v1_asd$NonVerbalIQ) 
dens (v1_asd$ADOS)
#Getting the priors
get_prior(m2, v1_asd, family = gaussian)
#Defining priors used
prior = c(prior(normal(0,0.5), class = 'Intercept'), prior(normal(0,0.5), class = 'b'), prior(normal(0,0.8), class = 'sigma'))
#modelling
m2_prior <-  brm(m2, family = gaussian, v1_asd, prior = prior, sample_prior = "only")
#Checking priors
pp_check(m0_prior, nsample = 100)
### MAP MAP MAP MAP
flist <-  alist(
    ADOS ~ dnorm(mu , sigma),
    mu <- a + b*NonVerbalIQ ,
        a ~ dnorm( 0 , 0.5 ) ,
        b ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dunif( 0 , 1 )
  )
m.m2 <- rethinking::map(flist, data = v1_asd)
precis(m.m2)
precis_plot(precis(m.m2))
```

```{r plots}
### Plot of raw data
ggplot(v1_asd, aes(NonVerbalIQ, ADOS)) +
  geom_point()
### Plot
# extract 20 samples from the posterior
post <- extract.samples( m.m2 , n=100 )
# display raw data and sample size
plot( ADOS ~ NonVerbalIQ , data=v1_asd )
abline( a=coef(m.m2)["a"] , b=coef(m.m2)["b"], col=col.alpha("red", 1))
for (i in 1:150)
abline( a=post$a[i], b=post$b[i], col=col.alpha("blue",0.1))
### new plot
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
nverb.seq <- seq(from= -3, to= 3, by=0.1)
# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu2 <- link( m.m2 , data=data.frame(NonVerbalIQ=nverb.seq) )
str(mu2)
###FUNKYPLOT 
# use type="n" to hide raw data
plot( ADOS ~ NonVerbalIQ , v1_asd , type="n" )
# loop over samples and plot each mu value
for ( i in 1:nrow(mu2) )
points( nverb.seq , mu2[i,] , pch=16 , col=col.alpha(rangi2,0.1) )
### GOODPPLOT NICEPLOT
# summarize the distribution of mu
mu2.mean <- apply(mu2 , 2 , mean )
mu2.HPDI <- apply(mu2 , 2 , HPDI , prob=0.89 )
# plot raw data
# fading out points to make line and interval more visible
plot( ADOS ~ NonVerbalIQ , data=v1_asd , col=col.alpha(rangi2,1) )
# plot the MAP line, aka the mean mu for each weight
lines( nverb.seq , mu2.mean )
# plot a shaded region for 89% HPDI
shade( mu2.HPDI , nverb.seq )
```


1.3. Social IQ and ADOS
```{r sIQ + ados}   
plot(v1_asd$ADOS ~ v1_asd$SocialIQ) 
 
#  defining model first
m3 <-  bf(ADOS ~ 1 + SocialIQ)
dens (v1_asd$SocialIQ)
dens (v1_asd$ADOS)
#Getting the priors 
get_prior(m3, v1_asd, family = gaussian)
#Defining priors used
prior = c(prior(normal(0,0.5), class = 'Intercept'), prior(normal(0,0.5), class = 'b'), prior(normal(0,0.8), class = 'sigma'))
#modelling
m3_prior <-  brm(m3, family = gaussian, v1_asd, prior = prior, sample_prior = "only")
#Checking priors
pp_check(m3_prior, nsample = 100)
#MAP
flist <-  alist(
    ADOS ~ dnorm(mu , sigma),
    mu <- a + b*SocialIQ ,
        a ~ dnorm( 0 , 0.5 ) ,
        b ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dunif( 0 , 1 )
  )
m.m3 <- rethinking::map(flist, data = v1_asd)
precis(m.m3)
precis_plot(precis(m.m3))
```

```{r plots}
### Plot of raw data 
ggplot(v1_asd, aes(SocialIQ, ADOS)) +
  geom_point()
### Plot
# extract 20 samples from the posterior
post <- extract.samples( m.m3 , n=100 )
# display raw data and sample size
plot( ADOS ~ SocialIQ , data=v1_asd )
abline( a=coef(m.m3)["a"] , b=coef(m.m3)["b"], col=col.alpha("red", 1))
for (i in 1:150)
abline( a=post$a[i], b=post$b[i], col=col.alpha("blue",0.1))
### new plot
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
soc.seq <- seq(from= -2, to= 3, by=0.1)
# use link to compute mu
Show in New WindowClear OutputExpand/Collapse Output
Modify Chunk OptionsRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output
 Show Traceback
Error in data.frame(SocialIQ = soc.seq) : object 'soc.seq' not found
# for each sample from posterior
# and for each weight in weight.seq
mu3 <- link( m.m3, data=data.frame(SocialIQ=soc.seq) )
str(m3)
###FUNKYPLOT
# use type="n" to hide raw data
plot( ADOS~ SocialIQ , v1_asd , type="n" )
# loop over samples and plot each mu value
for ( i in 1:nrow(mu3) ) 
points( soc.seq , mu3[i,] , pch=16 , col=col.alpha(rangi2,0.1) )
### GOODPPLOT NICEPLOT
# summarize the distribution of mu 
mu3.mean <- apply(mu3 , 2 , mean )
mu3.HPDI <- apply(mu3 , 2 , HPDI , prob=0.89 )
# plot raw data
# fading out points to make line and interval more visible
plot( ADOS ~ SocialIQ , data=v1_asd , col=col.alpha(rangi2,1) )
# plot the MAP line, aka the mean mu for each weight
lines( soc.seq , mu3.mean )
# plot a shaded region for 89% HPDI
shade( mu3.HPDI , soc.seq )
```


```{r residual check}
res_normal <- function(model, outcome, predictor) {
   
  mu <- coef(model)['a'] + 
    coef(model)['b']*predictor 
  
  m.resid <- outcome - mu
  dens(m.resid)
}
res_normal(m.m0, visit1$ADOS, visit1$VerbalIQ)
res_normal(m.m2, visit1$ADOS, visit1$NonVerbalIQ)
res_normal(m.m3, visit1$ADOS, visit1$SocialIQ)
```


### 2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
- Answered in docs

2.2. Build the model, assess its quality, write a few lines interpreting the results.
```{r model w all measures } 
#MAP
flist <-  alist(
    ADOS ~ dnorm(mu , sigma),
    mu <- a + bs*SocialIQ + bv*VerbalIQ + bn*NonVerbalIQ,
        a ~ dnorm( 0 , 0.5 ) ,
        bs ~ dnorm( 0 , 0.5 ) ,
        bv ~ dnorm( 0 , 0.5 ), 
        bn ~ dnorm(0, 0.5 ),
        sigma ~ dunif( 0 , 1 ) 
  )
m2.1 <- rethinking::map(flist, data = v1_asd)
precis(m2.1)
precis_plot(precis(m2.1))
#Extract sample from the model
post.samples=extract.samples(m2.1)
plot(bs ~ bn, post.samples, col = col.alpha(rangi2,0.1), pch = 16)
plot(bs ~ bv, post.samples, col = col.alpha(rangi2,0.1), pch = 16)
plot(bn ~ bv, post.samples, col = col.alpha(rangi2,0.1), pch = 16)
#plotting with the uncertainty
#does our model if the model makes sense?
plot(ADOS ~ SocialIQ + NonVerbalIQ + VerbalIQ, v1_asd)
abline(a=coef(m2.1)['a'], b=coef(m2.1)['bs']+coef(m2.1)['bn']+coef(m2.1)['bv'])
for ( i in 1:150 )
abline( a=post.samples$a[i] , b=post.samples$bs[i]+post.samples$bn[i]+post.samples$bv[i] , col=col.alpha(rangi2, 0.2) )
####5.11 i Rethinking
#Link-function to compute mu for each sample from posterior and for each value in the sequence
mu <- link(m2.1)
#summarize the across cases
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
#Plotting (5.12)
plot( mu.mean ~ v1_asd$ADOS , col=rangi2 , ylim=range(mu.HPDI) ,
xlab="Observed ADOS" , ylab="Predicted ADOS" )
abline( a=0 , b=1 , lty=2 )
for ( i in 1:nrow(v1_asd) )
lines( rep(v1_asd$ADOS[i],2) , c(mu.HPDI[1,i],mu.HPDI[2,i]) ,
col=rangi2 )
#5.13
#identify( x=V1_ASD$ADOS.s , y=mu.mean , labels=V1_ASD$ChildID , cex=0.8 )
```

```{r residual}
res_normal(m2.1, v1_asd$ADOS, v1_asd$VerbalIQ)
```


### 3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.
```{r all measures - asd+td} 
#Standardize predictors to a mean of 0 and sd of 1  
visit1 <- na.omit(visit1)
visit1$ADOS.s <-  (visit1$ADOS - mean(visit1$ADOS))/ sd(visit1$ADOS)
visit1$VerbalIQ.s <-  (visit1$VerbalIQ - mean(visit1$VerbalIQ, na.rm = T))/sd(visit1$VerbalIQ)
visit1$NonVerbalIQ.s <-  (visit1$NonVerbalIQ - mean(visit1$NonVerbalIQ, na.rm = T))/sd(visit1$NonVerbalIQ)
visit1$SocialIQ.s <-  (visit1$SocialIQ-mean(visit1$SocialIQ, na.rm = T))/sd(visit1$SocialIQ)
#Model
flist <-  alist(
    ADOS.s ~ dnorm(mu , sigma),
    mu <- a + bV*VerbalIQ.s + bN*NonVerbalIQ.s + bS*SocialIQ.s ,
        a ~ dnorm( 0 , 0.5 ) ,
        bV ~ dnorm( 0 , 0.5 ) ,
        bN ~ dnorm( 0 , 0.5 ) ,
        bS ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dunif( 0 , 0.8 )
  )
m3.1 <- rethinking::map(flist, data = visit1)
precis(m3.1)
precis_plot( precis(m3.1))
 
####5.11 i Rethinking
#Link-function to compute mu for each sample from posterior and for each value in the sequence
mu <- link(m3.1)
#summarize the across cases
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
#Plotting (5.12)
plot( mu.mean ~ visit1$ADOS.s , col= ('black') , ylim=range(mu.HPDI) ,
xlab="Observed ADOS" , ylab="Predicted ADOS" )
abline( a=0 , b=1 , lty=2 )
for ( i in 1:nrow(visit1) )
lines( rep(visit1$ADOS.s[i],2) , c(mu.HPDI[1,i],mu.HPDI[2,i]) ,
col=rangi2 )
#5.13
#identify( x=visit1$ADOS.s , y=mu.mean , labels=visit1$ChildID , cex=0.8 )
```

```{r interaction?}
#trying out interaction effect model. Doesn't work. abandon mission
flist <- 
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- intercept + gamma1*NonVerbalIQ.s + bgroup*ASD + gamma2*VerbalIQ.s + bgroup*ASD + gamma3*SocialIQ.s + bgroup*ASD,
   
    gamma1 <- bnon +  binteraction1*ASD,  #nonverbal
    gamma2 <- bverb + binteraction2*ASD,  #verbal
    gamma3 <- bsoc +  binteraction3*ASD,  #social
    
    intercept ~ dnorm(0, 0.5),
    
    bnon ~ dnorm(0, 0.5),
    bverb ~ dnorm(0, 0.5),
    bsoc ~ dnorm(0, 0.5),
    bgroup ~ dnorm(0,1),
    
    binteraction1 ~ dnorm(0,1),
    binteraction2 ~ dnorm(0,1),
    binteraction3 ~ dnorm(0,1),
    
    sigma ~ dnorm(0,0.8)
    )
    
m3.2 <- rethinking:: map(flist, data = visit1)
```

```{r residual}
res_normal(m3.1, v1_asd$ADOS, v1_asd$VerbalIQ)
```

```{r interaction?? plot}
visit1$ASD <- as.factor(visit1$ASD)
#checking with plots 
pacman::p_load(ggunchained)
p1 <- ggplot(visit1, aes(NonVerbalIQ, ADOS, color = ASD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "ADOS score by non-verbal IQ for ASD and TD") 
p2 <- ggplot(visit1, aes(VerbalIQ, ADOS, color = ASD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "ADOS score by verbal IQ for ASD and TD") 
p3 <- ggplot(visit1, aes(SocialIQ, ADOS, color = ASD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "ADOS score by social IQ for ASD and TD")
p1
p2
p3
```



### 4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?

```{r + Correlation + ASD only, TD only}
#filtering
V1_TD <- visit1 %>% filter(ASD == 0)
V1_TD <- (na.omit(V1_TD))
#standardising
V1_TD$ADOS.s = (V1_TD$ADOS - mean(V1_TD$ADOS))/ sd(V1_TD$ADOS)
V1_TD$VerbalIQ.s = (V1_TD$VerbalIQ - mean(V1_TD$VerbalIQ, na.rm = T))/sd(V1_TD$VerbalIQ)
V1_TD$NonVerbalIQ.s = (V1_TD$NonVerbalIQ - mean(V1_TD$NonVerbalIQ, na.rm = T))/sd(V1_TD$NonVerbalIQ)
V1_TD$SocialIQ.s = (V1_TD$SocialIQ-mean(V1_TD$SocialIQ, na.rm = T))/sd(V1_TD$SocialIQ)
#correlation
cor(select(v1_asd, VerbalIQ, NonVerbalIQ, SocialIQ))
cor(select(V1_TD, VerbalIQ, NonVerbalIQ, SocialIQ))
```

```{r + Residual ASD plots}
#Map 
flist <-  alist( 
    ADOS ~ dnorm(mu , sigma),
    mu <- a + bV*VerbalIQ + bN*NonVerbalIQ + bS*SocialIQ ,
        a ~ dnorm( 0 , 0.5 ) ,
        bV ~ dnorm( 0 , 0.5 ) ,
        bN ~ dnorm( 0 , 0.5 ) ,
        bS ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dunif( 0 , 0.8 )
  )
m1.4 <- rethinking::map(flist, data = v1_asd)
precis(m1.4)
precis_plot( precis(m1.4))
#Model for residuals of verbalIQ 'controlling' for NonVerbalIQ and SocialIQ
modelVIQ <- rethinking::map(
  alist(
    VerbalIQ ~ dnorm(mu, sigma),
    mu <- a + b1*NonVerbalIQ + b2*SocialIQ,
    a ~ dnorm(0,0.5),
    b1 ~ dnorm(0, 0.5),
    b2 ~ dnorm(0, 0.5),
    sigma ~ dunif(0, 0.8)
  ),
  data = v1_asd)
#creating model for residuals of non-verbalIQ 'controlling' for VIQ and SIQ
modelNonVIQ <- rethinking::map(
  alist(
    NonVerbalIQ ~ dnorm(mu, sigma),
    mu <- a + b1*VerbalIQ + b2*SocialIQ,
    a ~ dnorm(0,0.5),
    b1 ~ dnorm(0, 0.5),
    b2 ~ dnorm(0, 0.5),
    sigma ~ dunif(0, 0.8)
  ),
  data = v1_asd)
#creating model for residuals of social IQ 'controlling' for NVIQ and VIQ
modelSocIQ <- rethinking::map(
  alist(
    SocialIQ ~ dnorm(mu, sigma),
    mu <- a + b1*NonVerbalIQ + b2*VerbalIQ,
    a ~ dnorm(0, 0.5),
    b1 ~ dnorm(0, 0.5),
    b2 ~ dnorm(0, 0.5),
    sigma ~ dunif(0, 0.8)
  ),
  data = v1_asd)
#Residuals
#VERBAL
mu.v <- coef(modelVIQ)['a'] + coef(modelVIQ)['b1']*v1_asd$NonVerbalIQ + coef(modelVIQ)['b2']*v1_asd$SocialIQ
#compute residual for each value
resid_verbal <- v1_asd$VerbalIQ - mu.v
  
#SOCIAL
mu.s <- coef(modelSocIQ)['a'] + coef(modelSocIQ)['b1']*v1_asd$NonVerbalIQ + coef(modelSocIQ)['b2']*v1_asd$VerbalIQ
#compute residual for each value
resid_social <- v1_asd$SocialIQ - mu.s
  
#NONVERBAL
mu.n <- coef(modelNonVIQ)['a'] + coef(modelNonVIQ)['b1']*v1_asd$VerbalIQ + coef(modelNonVIQ)['b2']*v1_asd$SocialIQ
#compute residual for each value
resid_nonverbal <- v1_asd$NonVerbalIQ - mu.n
  
  
  
#Dataframes
VerbalIQ_df = data.frame(resid_verbal, v1_asd$ADOS)
NonVerbalIQ_df = data.frame(resid_nonverbal, v1_asd$ADOS)
SocialIQ_df = data.frame(resid_social, v1_asd$ADOS)
#### PLOTS PLOTS PLOTS
resplot_verbal = ggplot(VerbalIQ_df, aes(resid_verbal, v1_asd$ADOS)) +
  geom_point(aes(colour = palette(rainbow(34)))) +
  geom_smooth(method='lm', col = 'black') +
  geom_vline(xintercept=0, linetype = 'dotted') + theme_grey() + theme(legend.position = 'none')+
  xlab('Residuals for Verbal "controlling" for social and nonverbal')+
  ylab('Ados (Standardized)')
resplot_verbal
resplot_Nonverbal = ggplot(NonVerbalIQ_df, aes(resid_nonverbal, v1_asd$ADOS)) +
  geom_point(aes(colour = palette(rainbow(34)))) +
  geom_smooth(method='lm', col = 'black') +
  geom_vline(xintercept=0, linetype = 'dotted') + theme_grey() + theme(legend.position = 'none') + 
  xlab('Residuals for Nonverbal IQ "controlling" for social and verbal')+
  ylab('Ados (Standardized)')
resplot_social = ggplot(SocialIQ_df, aes(resid_social, v1_asd$ADOS)) +
  geom_point(aes(colour = palette(rainbow(34)))) +
  geom_smooth(method='lm', col = 'black') +
  geom_vline(xintercept=0, linetype = 'dotted') + theme_grey() + theme(legend.position = 'none') + 
  xlab('Residuals for Social IQ "controlling" for verbal and nonverbal')+
  ylab('Ados (Standardized)')
resplot_verbal
resplot_Nonverbal
resplot_social
```

```{r + Residual plots only TD}
#Model for residuals of verbalIQ 'controlling' for NonVerbalIQ and SocialIQ
modelVIQ <- rethinking::map(
  alist(
    VerbalIQ.s ~ dnorm(mu, sigma),
    mu <- a + b1*NonVerbalIQ.s + b2*SocialIQ.s,
    a ~ dnorm(0,0.5),
    b1 ~ dnorm(0, 0.5),
    b2 ~ dnorm(0, 0.5),
    sigma ~ dunif(0, 0.8)
  ),
  data = V1_TD)
#creating model for residuals of non-verbalIQ 'controlling' for VIQ and SIQ
modelNonVIQ <- rethinking::map(
  alist(
    NonVerbalIQ.s ~ dnorm(mu, sigma),
    mu <- a + b1*VerbalIQ.s + b2*SocialIQ.s,
    a ~ dnorm(0,0.5),
    b1 ~ dnorm(0, 0.5),
    b2 ~ dnorm(0, 0.5),
    sigma ~ dunif(0, 0.8)
  ),
  data = V1_TD)
#creating model for residuals of social IQ 'controlling' for NVIQ and VIQ
modelSocIQ <- rethinking::map(
  alist(
    SocialIQ.s ~ dnorm(mu, sigma),
    mu <- a + b1*NonVerbalIQ.s + b2*VerbalIQ.s,
    a ~ dnorm(0,0.5),
    b1 ~ dnorm(0, 0.5),
    b2 ~ dnorm(0, 0.5),
    sigma ~ dunif(0, 0.8)
  ),
  data = V1_TD)
#Residuals
#VERBAL
mu.v <- coef(modelVIQ)['a'] + coef(modelVIQ)['b1']*V1_TD$NonVerbalIQ.s + coef(modelVIQ)['b2']*V1_TD$SocialIQ.s
#compute residual for each value
resid_verbal <- V1_TD$VerbalIQ.s - mu.v
  
#SOCIAL
mu.s <- coef(modelSocIQ)['a'] + coef(modelSocIQ)['b1']*V1_TD$NonVerbalIQ.s + coef(modelSocIQ)['b2']*V1_TD$VerbalIQ.s
#compute residual for each value
resid_social <- V1_TD$SocialIQ.s - mu.s
  
#NONVERBAL
mu.n <- coef(modelNonVIQ)['a'] + coef(modelNonVIQ)['b1']*V1_TSD$VerbalIQ.s + coef(modelNonVIQ)['b2']*V1_TD$SocialIQ.s
#compute residual for each value
resid_nonverbal <- V1_TD$NonVerbalIQ.s - mu.n
  
  
  
#Dataframes
VerbalIQ_df = data.frame(resid_verbal, V1_TD$ADOS.s)
NonVerbalIQ_df = data.frame(resid_nonverbal, V1_TD$ADOS.s)
SocialIQ_df = data.frame(resid_social, V1_TD$ADOS.s)
resplot_verbal = ggplot(VerbalIQ_df, aes(resid_verbal, V1_TD.ADOS.s)) +
  geom_point(aes(colour = palette(rainbow(36)))) +
  geom_smooth(method='lm', col = 'black') +
  geom_vline(xintercept=0, linetype = 'dotted') + theme_grey() + theme(legend.position = 'none') + 
  xlab('Residuals for Verbal "controlling" for social and nonverbal')+
  ylab('Ados (Standardized)')
  
resplot_Nonverbal = ggplot(NonVerbalIQ_df, aes(resid_nonverbal, V1_TD.ADOS.s)) +
  geom_point(aes(colour = palette(rainbow(36)))) +
  geom_smooth(method='lm', col = 'black') +
  geom_vline(xintercept=0, linetype = 'dotted', col = 'black')+ theme_grey() + theme(legend.position = 'none') + 
xlab('Residuals for Nonverbal IQ "controlling" for social and verbal')+
  ylab('Ados (Standardized)')
resplot_social = ggplot(SocialIQ_df, aes(resid_social, V1_TD.ADOS.s)) +
  geom_point(aes(colour = palette(rainbow(36)))) +
  geom_smooth(method='lm', col = 'black') +
  geom_vline(xintercept=0, linetype = 'dotted')+ theme_grey() + theme(legend.position = 'none') + 
  xlab('Residuals for Social IQ "controlling" for verbal and nonverbal')+
  ylab('Ados (Standardized)')
resplot_verbal
resplot_Nonverbal
resplot_social
```

ASD children: 
Social IQ & Verbal IQ are the strongest predictors of ADOS scores
(redsidual plot)

Verbal IQ correlated with both Non-Verbal & Social IQ (more than 0.6)
Social & Non-Verbal IQ are not very correlated. 

TD children: 
Non-Verbal and Verbal IQ are highly correlated (0.55)
Social IQ is not really corrleated with any of them. 
Opposite effect on between social and nonverbal


4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.
